\documentclass[]{article}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{cleveref}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{amssymb}
\usetikzlibrary{calc,automata,positioning}
\def\istari{\textsc{Istari}\xspace}
\usepackage{float}
%opening
\input{defs}
\title{Data Plane Reactivity to Failures for WCMP}

\author{Kausik Subramanian, Loris D'Antoni, Aditya Akella}

\begin{document}

\maketitle

\begin{abstract}
We investigate data plane reactivity to WCMP under failures.
\end{abstract}

\section{Introduction}
In WCMP, a switch splits traffic for a destination based on 
configured weights (which is implemented by multiple 
forwarding rules for the same port and using a hashing function
to choose the ports). 
These weights are computed by a centralized controller based
on oversubscription and switch table sizes.

\paragraph{Failures.} When a link failure occurs,
the switch will simply split the traffic amongst the 
active links in the configured weights as before, this
can lead to greater imbalance of traffic along certain 
links. Between the time of failure and time taken by 
the WCMP controller to configure new weights, the 
network throughput can be deteriorated (can perform simple
simulations to measure this).

\section{Problem Statement}
In this project, we wish to investigate local data plane
reactivity mechanisms to failures. For instance, when a 
switch sees a link failure, if it can modify its weights 
locally (this however will not necessarily be optimal), we
could increase the network throughput more than under the 
failure scenario. 

Only modifying the weights of two switches (ends of the failed link)
may not yield good returns. We could investigate changes to switches
$n-$hops from the failed link for better reactivity to the failure. 

\paragraph{Output.} Synthesize data-plane strategies
for switches to react locally to link failures for WCMP. We can 
consider a network of P4 switches for this purpose. Strategy-effectiveness
 can be measured by different metrics - network throughput, number of 
 congested links etc. 



\end{document}